Hydras, Ordinals \& Co.\\ A library in Coq of entertaining formal mathematics

### 2021-11-12 doc/hydras.tex
small changes

\index{coq}{Commands!Print Assumptions}

@@

% \begin{todo}
%  Still very incomplete!
% \end{todo}
{\Large \textbf{In progress} This index is currently under reorganization for a few days. We apologize for its incompleteness! }

\printindex{coq}{Coq, plug-ins and standard library}
\printindex{maths}{Mathematical notions and algorithmics}
\printindex{hydras}{Library hydras: Ordinals and hydra battles}
\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
\printindex{additions}{Library additions: Addition chains}

\section{Main notations}

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Ordinals and ordinal notations}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\

F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
\pageref{sect:notation-FS}\\ 
omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\

\hline
\end{tabular}
\begin{tablenotes}
  \item[1] This order is total, but not well-founded, because of not well formed terms.
\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
\item[3] This order is total \emph{and} well-founded.
\item [4]
Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
  \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal. Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}[H]

  \begin{threeparttable}
    \caption{hydra battles}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
\hline
\end{tabular}

%\begin{tablenotes}
%\end{tablenotes}

  \end{threeparttable}

\end{table}

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Addition chains}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
\hline
\end{tabular}

%\begin{tablenotes}
%\end{tablenotes}

  \end{threeparttable}

\end{table}

>>>>

%\index{coq}{Commands!Print Assumptions}

@@

% \begin{todo}
%  Still very incomplete!
% \end{todo}
{\Large \textbf{In progress} This index is currently under reorganization.  We apologize for its incompleteness! }

\printindex{coq}{Coq, plug-ins and standard library}
\printindex{maths}{Mathematical notions and algorithmics}
%\printindex{hydras}{Library hydras: Ordinals and hydra battles}
%\printindex{primrec}{Library hydras.Ackermann: Primitive recursive functions}
%\printindex{additions}{Library additions: Addition chains}

% \section{Main notations}

% \begin{table}[H]
%   \centering
%   \begin{threeparttable}
%     \caption{Ordinals and ordinal notations}
% \begin{tabular}{|r | c|c|c|c|l|}
% \hline
% Name & Gallina&Math& Description& Page \\\hline
% \texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
% \texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
% \texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
% \texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
%  \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{4} & \pageref{sect:on-lt-notation}\\
%  \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
% \texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
% \texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\

% F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
% \pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
% FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{5} &  
% \pageref{sect:notation-FS}\\ 
% omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
% phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\

% \hline
% \end{tabular}
% \begin{tablenotes}
%   \item[1] This order is total, but not well-founded, because of not well formed terms.
% \item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
% \item[3] This order is total \emph{and} well-founded.
% \item [4]
% Some notations may belong to several scopes. For instance, ``\texttt{o<}'' is
% bound in \texttt{ON\_scope}, \texttt{E0\_scope}, \texttt{t1\_scope}, etc., and locally in several libraries.
%   \item [5] Note that there exist also various coercions from \texttt{nat} to types of ordinal. Depending on the current scope and  \coq's syntactic analysis algorithm, \texttt{F} may be left implicit.
% \end{tablenotes}
% \end{threeparttable}
% \end{table}

% \vspace{4pt}

% \begin{table}[H]

%   \begin{threeparttable}
%     \caption{hydra battles}
% \begin{tabular}{|c|c|c|c|l|}
% \hline
% Name & Gallina&Math& Description& Page \\\hline
% \texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
% \texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
% \texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\
% \hline
% \end{tabular}

% %\begin{tablenotes}
% %\end{tablenotes}

%   \end{threeparttable}

% \end{table}

% \begin{table}[H]
%   \centering
%   \begin{threeparttable}
%     \caption{Addition chains}
% \begin{tabular}{|c|c|c|c|l|}
% \hline
% Name & Gallina&Math& Description& Page \\\hline
% \texttt{Mult} & \texttt{$z$ <--- $x$ times $y$} & & monadic notation & \pageref{monadic-mult} \\
% \hline
% \end{tabular}

% %\begin{tablenotes}
% %\end{tablenotes}

%   \end{threeparttable}

% \end{table}

### 2021-11-12 doc/part-hydras.tex
minor changes
small changes

\index{coq}{Mutually inductive types}
\index{coq}{Commands!Scheme}

@@

\subsection{A  small proof of impossibility}
\index{coq}{Proofs of impossibility}

@@

\index{coq}{Commands!Function}

@@

\index{coq}{Commands!Program}

@@

\subsubsection{A first attempt}
\label{sec:orgheadline77}
\index{coq}{Well-founded induction}

@@

%ici
    \section{A variant for hydra battles}

@@

\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
The type \texttt{fuel}, an alternative 
to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
\index{coq}{Giving fuel to a long computation}

@@

\label{lemma:L-2_6-1}
\index{coq}{Sigma types}

>>>>

\index{coq}{Mutually inductive types}
%\index{coq}{Commands!Scheme}

@@

\subsection{A  small proof of impossibility}
%\index{coq}{Proofs of impossibility}

@@

%\index{coq}{Commands!Function}

@@

%\index{coq}{Commands!Program}

@@

\subsubsection{A first attempt}
\label{sec:orgheadline77}
%\index{coq}{Well-founded induction}

@@

\begin{exercise}[not trivial]
  It may be interesting to write a \emph{direct} proof  of well-foundedness of the order in $\omega^\omega$ (\emph{i.e.}
  without using properties of $\epsilon_0$). This exercise may
  help to understand better the proof structure of Sect.~\vref{sec:strongly-accessible}.
    \end{exercise}
%ici
    \section{A variant for hydra battles}

@@

\textbf{Hint:} You may add to your function a third argument usually called \texttt{fuel} for allowing you to give a structurally 
recursive function (\emph{cf} the post of Guillaume Melquiond on Coq-club (Dec 21, 2020)
\url{https://sympa.inria.fr/sympa/arc/coq-club/2020-12/msg00069.html}).
The type \texttt{fuel}, an alternative 
to \texttt{nat} is available on \href{../theories/html/hydras.Prelude.Fuel.html}{Prelude.Fuel}).
%\index{coq}{Giving fuel to a long computation}

@@

\label{lemma:L-2_6-1}
%\index{coq}{Sigma types}
